Kernel Development Using Rust: A Systematic Literature Review

### 2024-05-24 paper.tex
Update on Overleaf.

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>00000000.0000000.0000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc[300]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc[100]{Do Not Use This Code~Generate the Correct Terms for Your Paper}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Memory safety}

\received{20 February 2007}
\received[revised]{12 March 2009}
\received[accepted]{5 June 2009}

@@

Once we had our initial collection of papers from the databases we merged all duplicate papers into one record and then started our first pass which consisted of reading the title and abstract and either marking the paper as \textit{include} or \textit{exclude}. Once  the initial screening was completed we were left with 29 potential papers that we needed an in depth reading.

@@

Liang et al.~\cite{Liang2021-bo} developed a new operating system named Rustpi with a micro-kernel architecture implemented in Rust to explore how modern language features can help with building reilabe systems.

\begin{table}
    \begin{tabular}{||l|l|l||}
    \hline
    Approach & Papers & Operating System in Rust\\
    \hline\hline
    Monolithic  & \cite{The_kernel_development_community_undated-iw}, \cite{Li2019-ru} & \href{https://docs.kernel.org/rust/}{Linux kernel v6.1+}\\
    Micro-kernel & \cite{Chen2023-wb},\cite{Liang2021-bo}, \cite{Liu2024-xe} & Atmosphere, Redox, Redleaf\\
    Embedded & \cite{Culic2022-bk} & \href{https://github.com/tock/tock}{Tock}, \href{https://hubris.oxide.computer/}{Hubris}, \href{https://www.drone-os.com/}{Drone}, \href{https://bern-rtos.org/}{Bern} \\
    Unikernel & \cite{Lankes2019-cm},  \cite{Boos2020-zh}, \cite{Ijaz2023-da}  & RustyHermit, Theseus \\
    Exokernel & \cite{Li2024-yb} & W-Kernel \\
    \hline
  \end{tabular}
  \caption{Approaches and Methodologies for Rust in the Kernel}
    \label{tab:RQ1}
\end{table}

@@

            \item Support typed assembly language for Rust
            \item Support trusted build environments 
            \item Provide software-only stack guard with extensible probing interface
            \item Develop zero-copy serialization of “plain- old” data structures
        \end{itemize}
        \\
        \hline
        3 & Soundness~\cite{Klimt2023-ob} &
        \begin{itemize}
            \item Unsynchronized Global State - any use of mutable statics is unsafe
            \item C-Style Abstractions - Use rust style abstractions to properly encapsulate internal unsafe usage of raw pointers
            \item Aliasing Mutable References - Giving out raw pointers to memory that is also referenced mutably
            \item Re-implementing Memory Access - instead of accessing specific memory regions in assembly, creating and using references to the whole region is preferable
        \end{itemize}
        \\
        \hline
        4 & Panics \cite{Ma2023-ef},\cite{Burtsev2021-mh} &
        \begin{itemize}
            \item Support extendable, no\_std unwind library
            \item Stack unwinding in embedded environments
        \end{itemize}
        \\
    \hline
  \end{tabular}
  \caption{Challenges Unique to the Rust Programming Language}
    \label{tab:RQ3}
\end{table*}

>>>>

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002944.10011122.10002945</concept_id>
<concept_desc>General and reference~Surveys and overviews</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10011006.10011008.10011024</concept_id>
<concept_desc>Software and its engineering~Language features</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10010520.10010521</concept_id>
<concept_desc>Computer systems organization~Architectures</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{General and reference~Surveys and overviews}
\ccsdesc[500]{Software and its engineering~Language features}
\ccsdesc[500]{Computer systems organization~Architectures}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Memory safety, Rust, Kernel, Operating System}

\received{30 May 2024}
\received[revised]{TBD}
\received[accepted]{TBD}

@@

Once we had our initial collection of papers from the databases we merged all duplicate papers into one record and then started our first pass which consisted of reading the title and abstract and either marking the paper as \textit{include} or \textit{exclude}. Once  the initial screening was completed we were left with 28 potential papers that we needed an in depth reading.

@@

Liang et al.~\cite{Liang2021-bo} developed a new operating system named Rustpi with a micro-kernel architecture implemented in Rust to explore how modern language features can help with building reliable systems.

\begin{table}
    \begin{tabular}{||l|l|l||}
    \hline
    Approach & Papers & Operating System in Rust\\
    \hline\hline
    Monolithic  & \cite{The_kernel_development_community_undated-iw}, \cite{Li2019-ru}, \cite{Miller2021-pg} & \href{https://docs.kernel.org/rust/}{Linux kernel v6.1+}\\
    Micro-kernel & \cite{Chen2023-wb},\cite{Liang2021-bo}, \cite{Liu2024-xe} \cite{} & Atmosphere, Redox, Redleaf\\
    Embedded & \cite{Culic2022-bk} & \href{https://github.com/tock/tock}{Tock}, \href{https://hubris.oxide.computer/}{Hubris}, \href{https://www.drone-os.com/}{Drone}, \href{https://bern-rtos.org/}{Bern} \\
    Unikernel & \cite{Lankes2019-cm},  \cite{Boos2020-zh}, \cite{Ijaz2023-da}  & RustyHermit, Theseus \\
    Exokernel & \cite{Li2024-yb} & W-Kernel \\
    \hline
  \end{tabular}
  \caption{Approaches and Methodologies for Rust in the Kernel}
    \label{tab:RQ1}
\end{table}

@@

            \item Support typed assembly language for Rust
            \item Support trusted build environments 
            \item Provide software-only stack guard with extensible probing interface
            \item Develop zero-copy serialization of “plain- old” data structures
        \end{itemize}
        \\
        \hline
        3 & Soundness~\cite{Klimt2023-ob} &
        \begin{itemize}
            \item Unsynchronized Global State - any use of mutable statics is unsafe
            \item C-Style Abstractions - Use rust style abstractions to properly encapsulate internal unsafe usage of raw pointers
            \item Aliasing Mutable References - Giving out raw pointers to memory that is also referenced mutably
            \item Re-implementing Memory Access - instead of accessing specific memory regions in assembly, creating and using references to the whole region is preferable
        \end{itemize}
        \\
        \hline
        4 & Panics \cite{Ma2023-ef},\cite{Burtsev2021-mh} &
        \begin{itemize}
            \item Support extendable, no\_std unwind library
            \item Stack unwinding in embedded environments
        \end{itemize}
        \\
        \hline
        5 & C Interop \cite{Miller2019-xm} &
        \begin{itemize}
            \item Kernel interfaces, while designed for extensibility, are not designed for type safety
            \item Hybrid Code Flow. The rust compiler can not track ownership when switching between modules written in C and Rust
        \end{itemize}
        \\
    \hline
  \end{tabular}
  \caption{Challenges Unique to the Rust Programming Language}
    \label{tab:RQ3}
\end{table*}
